---
title: "Reproducible research project 1"
author: "Yu-Fen (Andrea) Wang"
date: "October 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load the data 
```{r}
url="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,destfile = "./project1.zip")
## upzip the file
unzip(zipfile = "./project1.zip",exdir=getwd())
## open the data file
data<-read.csv(file = "activity.csv",header=T,sep=",")
```

## exploratory data analysis
```{r}
## let's take a look at what data looks like
str(data)

```
Now we know the variables are:
- steps: integer,missing value presented as NA
- date: factor, presented as YYYY-MM-DD
- interval: integer, increase by 5, it shows which interval the steps are measured 

## preprocess data
```{r}
## change the date column into date formart
library(lubridate)
data$date<-ymd(data$date)
head(data,5)

## remove those rows without step value
data2<-data[!is.na(data$steps),]

summary(data2)
```

## calculate numbers of steps taken per day
```{r}
steps_day<-aggregate(data2$steps~data2$date,FUN=sum)
names(steps_day)<-c("Date","Steps")
head(steps_day,5)
```

## Make Histogram for number of steps taken per day

```{r}

hist(steps_day$Steps,xlab="Steps",main="Steps Taken per Day",labels = T)

```
## Mean and median number of steps taken each day
```{r}
str(steps_day)
mean(steps_day$Steps)
median(steps_day$Steps)
```
The mean steps taken each day is 10766 steps
The median steps taken each day is 10765 steps

## Time series plot of the average number of steps taken
```{r}
## calculate the avg steps taken per interval
steps_interval<-aggregate(data2$steps~data2$interval,FUN=sum)
names(steps_interval)<-c("interval","steps")
library(ggplot2)
p<-ggplot(steps_interval,aes(x=interval,y=steps))
p+geom_line(color="blue",size=1)+xlab("Interval")+ylab("Steps")+ggtitle("Average Steps Taken Per Interval")
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max(steps_interval$steps)
steps_interval[steps_interval$steps==max(steps_interval$steps),]
```
Thte interval with max steps is 835

##Code to describe and show a strategy for imputing missing data
```{r}
sum(is.na(data$steps))

```
There are 2304 rows without steps value

## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
na<-data[is.na(data$steps),]
summary(na)
## in order to show missing value in new graph, I assign them with a desinated value as mean steps 100
na$steps=100
steps_day_na<-aggregate(na$steps~na$date,FUN=sum)
names(steps_day_na)=c("Date","Steps")
head(steps_day_na,5)
## then I merge the modified missing value set to steps_day
merge_na<-rbind(steps_day,steps_day_na)
## plot two histogram with different color to illustrate the difference
hist(merge_na$Steps,xlab="Steps",main="Steps Taken per Day excluding NA Values",col="light blue",labels = T)
hist(steps_day$Steps,xlab="Steps",main="Steps Taken per Day",col="light green",add=T)

legend("topright",c("No-NA","With NA data"),fill = c("light green","light blue"))
```
## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(dplyr)
## create a new colum to convert the date into weekday
data2<-mutate(data2,day=wday(data2$date,label = T))
head(data2,5)
## create a column to convert the day into weekday or weekend

weekend<-data2[data2$day=="Sat"|data2$day=="Sun",]
weekend<-mutate(weekend,wday="weekend")
weekday<-anti_join(data2,weekend,by='day')
weekday<-mutate(weekday,wday="weekday")
data3<-rbind(weekend,weekday)
## summarize steps by interval and wday
steps_interval_2<-aggregate(data3$steps~data3$interval+data3$wday,FUN=mean)
names(steps_interval_2)<-c("interval","wday","steps")

## make 2 plot to show the average steps by interval in weekday vs weekend
library(ggplot2)
qplot(interval,steps,data=steps_interval_2,facets = wday~.,geom = "line")+ggtitle("Average steps per interval: weekday vs weekends")

```


